`r params$doc_title`

### 2022-11-20 scripts/MSc.Rmd
performance plots
ggarrange plot

plot_data1b <- df_cum_expl_var %>% 
  pivot_longer(cols = c(cum_expl_varX2, cum_expl_varU, cum_expl_varY)) %>% 
  mutate(name = factor(name, labels = c("% var(U)", "% var(X)", "% var(Y)")),
         name = factor(name, levels = c("% var(X)", "% var(Y)", "% var(U)") )
         )
## plot
ggplot(plot_data1b, aes(x=n, y=value, col=name))+
  geom_point(alpha=0.6)+
  geom_line(alpha=0.6)+
  geom_line(aes(x=n, y=count_IVs_scaled, col="Number of IVs"), linetype="dotdash")+
  scale_x_continuous(name = paste0("Size of sample ", as.roman(1)),  # 2-sample-MR or label "Study Size"
                     labels = comma, 
                     breaks = N_annotation, #c( 10000, n[7:length(n)]),
                     limits = c(10000, 300000))+
  #ylab("% variance explained by IVs")+
  labs(title = "Figure 1",
       col='', 
       caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
       ) +
  scale_y_continuous(
    # Features of the first axis
    name = "% variance explained by IVs",
    # Add a second axis and specify its features
    sec.axis = sec_axis(~./scale_factor, name="Number of IVs")
  ) + 
  theme_few()+
  #scale_colour_few() +
  #scale_color_manual(name = "", values = c("% var(U)" = "red", 
  #                                         "% var(X)" = "blue", 
  #                                         "% var(Y)" = "darkblue", 
  #                                         "Number of IVs" = "black"
  #                                         ))
  #
  scale_color_manual(name = "", values = c("% var(U)" = ggthemes_data$few$colors$Dark[2,2][[1]], 
                                           "% var(X)" = ggthemes_data$few$colors$Dark[3,2][[1]], 
                                           "% var(Y)" = ggthemes_data$few$colors$Dark[4,2][[1]], 
                                           "Number of IVs" = "black"),
                     # order
                     breaks = c( "Number of IVs", "% var(X)", "% var(Y)", "% var(U)")
  ) -> fig01b.update
fig01b.update

@@

  N_annotation <- c(10000, 100000, 200000, 300000)
  ## plot
  plotsubtitle <- paste0("Relatively ", maf ," variants\nwith relatively ", effect_size, " effects")
  ggplot(plot_data1, aes(x=n, y=value, col=name))+
    geom_point(alpha=0.6)+
    geom_line(alpha=0.6)+
    geom_line(aes(x=n, y=count_IVs_scaled, col="Number of IVs"), linetype="dotdash")+
    scale_x_continuous(name = paste0("Size of sample ", as.roman(1)),  # 2-sample-MR or label "Study Size"
                       labels = comma, 
                       breaks = N_annotation, #c( 10000, n[7:length(n)]),
                       limits = c(10000, 300000))+
    #ylab("% variance explained by IVs")+
    labs(title = "Figure 1", subtitle = plotsubtitle,
         col='', 
         caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits   explained by the IVs"
         ) +
    scale_y_continuous(
      # Features of the first axis
      name = "% variance explained by IVs",
      # Add a second axis and specify its features
      sec.axis = sec_axis(~./scale_factor, name="Number of IVs")
    ) + 
    theme_few()+
    #scale_colour_few() +
    #scale_color_manual(name = "", values = c("% var(U)" = "red", 
    #                                         "% var(X)" = "blue", 
    #                                         "% var(Y)" = "darkblue", 
    #                                         "Number of IVs" = "black"
    #                                         ))
    #
    scale_color_manual(name = "", values = c("% var(U)" = ggthemes_data$few$colors$Dark[2,2][[1]], 
                                             "% var(X)" = ggthemes_data$few$colors$Dark[3,2][[1]], 
                                             "% var(Y)" = ggthemes_data$few$colors$Dark[4,2][[1]], 
                                             "Number of IVs" = "black"),
                       # order
                       breaks = c( "Number of IVs", "% var(X)", "% var(Y)", "% var(U)")
    ) -> figure01.quadrant
  print(figure01.quadrant)

  filenamefig01_quadrant <- paste0("../output/plots/", Sys.Date(), "_figure01_", 
                                   "EFFECT_SIZE_", effect_size,
                                   "_VARIANTS_", maf,
                                   ".png")
  ggsave(filename = filenamefig01_quadrant, plot = figure01.quadrant, 
         width = 10, height = 6, 
         units = "in"  # default
         )
}

@@

@@

## START SIM
T0 = proc.time()[3]
for (i in 1:length(colnames_pval)) {
 threshold <- 5*10^-6
 se_update_BMI <- c()
 se_update_GBC <- c()

@@

saveRDS(est_sim, paste0("../output/Rdata/", Sys.Date(), "_est_sim.rds"))
est_sim <- readRDS(file = "../output/Rdata/2022-11-15_est_sim.rds")
```

@@

```{r plot-performance-metrics}
#BIAS
ggplot(performance_metrics, aes(x=N, y=bias))+ 
  geom_point()

#MSE
ggplot(performance_metrics, aes(x=N, y=mse))+ 
  geom_point()

# compare distribution of simulated thetas against theta_324000
boxplot_dat <- est_sim %>% 
  select(!!!(colname_theta)) %>% 
  pivot_longer(cols = everything(), names_to = "theta_sim")

ggplot(boxplot_dat, aes(x=theta_sim, y=value))+
  geom_boxplot()+
  coord_flip()

>>>>

plot_data1b <- df_cum_expl_var %>% 
  pivot_longer(cols = c(cum_expl_varX2, cum_expl_varU, cum_expl_varY)) %>% 
  mutate(name = factor(name, labels = c("% var(U)", "% var(X)", "% var(Y)")),
         name = factor(name, levels = c("% var(X)", "% var(Y)", "% var(U)") )
         )
## plot
ggplot(plot_data1b, aes(x=n, y=value, col=name))+
  geom_point(alpha=0.6)+
  geom_line(alpha=0.6)+
  geom_line(aes(x=n, y=count_IVs_scaled, col="Number of IVs"), linetype="dotdash")+
  scale_x_continuous(name = paste0("Size of sample ", as.roman(1)),  # 2-sample-MR or label "Study Size"
                     labels = comma, 
                     breaks = N_annotation, #c( 10000, n[7:length(n)]),
                     limits = c(10000, 300000))+
  #ylab("% variance explained by IVs")+
  labs(title = "Figure 1",
       col='', 
       caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
       ) +
  scale_y_continuous(
    # Features of the first axis
    name = "% variance explained by IVs",
    # Add a second axis and specify its features
    sec.axis = sec_axis(~./scale_factor, name="Number of IVs"),
    limits = c(0,5)
  ) + 
  theme_few()+
  #scale_colour_few() +
  #scale_color_manual(name = "", values = c("% var(U)" = "red", 
  #                                         "% var(X)" = "blue", 
  #                                         "% var(Y)" = "darkblue", 
  #                                         "Number of IVs" = "black"
  #                                         ))
  #
  scale_color_manual(name = "", values = c("% var(U)" = ggthemes_data$few$colors$Dark[2,2][[1]], 
                                           "% var(X)" = ggthemes_data$few$colors$Dark[3,2][[1]], 
                                           "% var(Y)" = ggthemes_data$few$colors$Dark[4,2][[1]], 
                                           "Number of IVs" = "black"),
                     # order
                     breaks = c( "Number of IVs", "% var(X)", "% var(Y)", "% var(U)")
  ) -> fig01b.update
fig01b.update

@@

  N_annotation <- c(10000, 100000, 200000, 300000)
  ## plot
  plotsubtitle <- paste0("Relatively ", maf ," variants\nwith relatively ", effect_size, " effects")
  ggplot(plot_data1, aes(x=n, y=value, col=name))+
    geom_point(alpha=0.6)+
    geom_line(alpha=0.6)+
    geom_line(aes(x=n, y=count_IVs_scaled, col="Number of IVs"), linetype="dotdash")+
    scale_x_continuous(name = paste0("Size of sample ", as.roman(1)),  # 2-sample-MR or label "Study Size"
                       labels = comma, 
                       breaks = N_annotation, #c( 10000, n[7:length(n)]),
                       limits = c(10000, 300000))+
    #ylab("% variance explained by IVs")+
    labs(title = "Figure 1", subtitle = plotsubtitle,
         col='', 
         caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits   explained by the IVs"
         ) +
    scale_y_continuous(
      # Features of the first axis
      name = "% variance explained by IVs",
      # Add a second axis and specify its features
      sec.axis = sec_axis(~./scale_factor, name="Number of IVs"),
      limits = c(0,2)
    ) + 
    theme_few()+
    #scale_colour_few() +
    #scale_color_manual(name = "", values = c("% var(U)" = "red", 
    #                                         "% var(X)" = "blue", 
    #                                         "% var(Y)" = "darkblue", 
    #                                         "Number of IVs" = "black"
    #                                         ))
    #
    scale_color_manual(name = "", values = c("% var(U)" = ggthemes_data$few$colors$Dark[2,2][[1]], 
                                             "% var(X)" = ggthemes_data$few$colors$Dark[3,2][[1]], 
                                             "% var(Y)" = ggthemes_data$few$colors$Dark[4,2][[1]], 
                                             "Number of IVs" = "black"),
                       # order
                       breaks = c( "Number of IVs", "% var(X)", "% var(Y)", "% var(U)")
    ) -> figure01.quadrant
  print(figure01.quadrant)

  filenamefig01_quadrant <- paste0("../output/plots/", Sys.Date(), "_figure01_", 
                                   "EFFECT_SIZE_", effect_size,
                                   "_VARIANTS_", maf,
                                   ".png")
  ggsave(filename = filenamefig01_quadrant, plot = figure01.quadrant, 
         width = 10, height = 6, 
         units = "in"  # default
         )

  # store plot in global environment
  plotname <<- paste0("figure01_EFFECT_SIZE_", effect_size, "_VARIANTS_", maf)
  assign(plotname, figure01.quadrant, envir = .GlobalEnv)
  #list2env(assign(plotname, figure01.quadrant), envir = .GlobalEnv)

@@

## FINAL PLOT

```{r fig.height=10, fig.width=14, warning=FALSE}
ggpubr::ggarrange(fig01b.update, fig02.update, 
                  figure01_EFFECT_SIZE_strong_VARIANTS_common, figure01_EFFECT_SIZE_strong_VARIANTS_rare,
                  figure01_EFFECT_SIZE_weak_VARIANTS_common, figure01_EFFECT_SIZE_weak_VARIANTS_rare, 
                  ncol=2, nrow=3)
```

@@

runSimulation <- FALSE

## START SIM
if (runSimulation==TRUE) {
T0 = proc.time()[3]
for (i in 1:length(colnames_pval)) {
 threshold <- 5*10^-6
 se_update_BMI <- c()
 se_update_GBC <- c()

@@

saveRDS(est_sim, paste0("../output/Rdata/", Sys.Date(), "_est_sim.rds"))
}

# read dat
est_sim <- readRDS(file = "../output/Rdata/2022-11-18_est_sim.rds")
```

@@

### Performance plots

```{r plot-performance-metrics}
#BIAS
ggplot(performance_metrics, aes(x=N, y=bias))+ 
  geom_point()+
  scale_x_continuous(labels = comma)+
  theme_few()

#MSE
ggplot(performance_metrics, aes(x=N, y=mse))+ 
  geom_point()+
  scale_x_continuous(labels = comma)+
  theme_few()

# compare distribution of simulated thetas against theta_324000
boxplot_dat <- est_sim %>% 
  select(!!!(colname_theta)) %>% 
  pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
  mutate(theta_sim_numeric = format(as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")), scientific=FALSE))

ggplot(boxplot_dat, aes(x=forcats::fct_rev( as.factor(theta_sim_numeric)), y=value)) +
  geom_boxplot()+
  geom_jitter(width=0.1,alpha=0.2)+
  #scale_x_discrete(name = paste0("Size of sample ", as.roman(1)),  breaks = c(" 50000", "100000", "200000", "300000"), )+ 
  ylab(expression(paste(theta[sim], " with ", n[sim]== 25 ) )  ) +
  labs(title = "Estimates from beta simulation",
       caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
       )+
  coord_flip()+
  theme_few()
```
